---
title: "Exam-02"
author: "Pablo Vivas"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-location: right
    toc-depth: 1
    code-copy: true
    code-fold: true
    highlight-style: nord
  docx:
    toc: true
    toc-depth: 3
    fig-align: center
    df-print: tibble
    highlight-style: nord
    output-file: exam-02-pablo-vivas.docx
execute:
  warning: false
  error: false
  echo: true
theme: simplex
---

```{r}
#| label: set-up
#| echo: false

library(broom)
library(gt)
library(lavaan)
library(psych)
library(stringr)
library(tidyverse)
```

# Question 1

```{r}
#| label: read data

data = read_csv("./data/personality-1.csv")

data_sub = data |> 
  select(str_c("pers", c("03", "07", 11, 13, 16, 26, 36)))
```

## a)

The single-factor model using seven indicators is statistically identified. This because the model contains $14$ parameters $(\lambda_1,...,\lambda_7, \varepsilon_1, ..., \varepsilon_7)$ and the correlation matrix contains $\frac{7(7+1)}{2} = \frac{56}{2} = 28$ unique pieces of information, also known as degrees of freedom. Hence, we have more degrees of freedom than model parameters which results in the model being statistically identified. (We have $28-14 = 14$ degrees of freedom to spare 🥳)

## b)

```{r}
#| label: tbl-fit-one-factor-cfa
#| tbl-cap: Fit measures for one-factor model 

mod_01 = '
  f1 =~ pers03 + pers07 + pers11 + pers13 + pers16 + pers26 + pers36
'

cfa_01 = cfa(mod_01, data = data_sub)

fitMeasures(cfa_01) |> 
  tidy() |>
  mutate(
    measure = names,
    value = round(x, 2),
    .keep = "unused") |> 
  filter(measure %in% c("chisq", "pvalue", "cfi", "tli", "rmsea", "srmr")) |> 
  gt() |> 
  cols_width(
    value ~  px(100)
  )
```

Given the results of @tbl-fit-one-factor-cfa, the model doesn't seem to fit the data. 

## c)

```{r}
residuals(cfa_01, type = "standardized")
```

```{r}
modindices(cfa_01, sort = T)
```

## d)

```{r}
mod_02 = '
  f1 =~ pers03 + pers07 + pers11 + pers13 + pers16 + pers26 + pers36
  pers16	~~	pers26
  pers11	~~	pers16
  pers16	~~	pers36
  pers26	~~	pers36
  pers11	~~	pers26
  pers03	~~	pers13
  pers11	~~	pers36
  pers13	~~	pers16
  pers13	~~	pers26
  pers03	~~	pers36
'

cfa_02 = cfa(mod_02, data = data_sub)

fitMeasures(cfa_02) |> 
  tidy() |>
  mutate(
    measure = names,
    value = round(x, 2),
    .keep = "unused") |> 
  filter(measure %in% c("chisq", "pvalue", "cfi", "tli", "rmsea", "srmr")) |> 
  gt() |> 
  cols_width(
    value ~  px(100)
  )
```

```{r}
anova(cfa_01, cfa_02)
```

## e)

# Question 2

## a)

```{r}
cor_mat = data |> 
  drop_na() |> 
  cor()

cortest.bartlett(cor_mat, n = nrow(data |> drop_na()))

KMO(data |> drop_na())
```

## b)

```{r}
fa.parallel(data, fm = "ml", n.obs = nrow(data |> drop_na()))

scree(data |> drop_na())

efa_01 = fa(cor_mat, 
            nfactors = 6, 
            n.obs = nrow(data |> drop_na()), 
            fm="ml", 
            rotate = "none")
```

## c)

```{r}
efa_02 = fa(cor_mat, 
            nfactors = 6, 
            n.obs = nrow(data |> drop_na()), 
            fm="ml", 
            rotate = "promax")
```


## d)

## e)

## f)

# Question 3

## a)

## b)
